An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-11 draft.tex
some references added

In the last decades, nature-inspired optimization algorithms have been succesfully
applied to solve complex real-world problems \cite{yang2014nature}. Algorithms
inspired by natural processes include evolutionary algorithms (EAs)
\cite{back1996evolutionary} and swarm intelligence (SI) \cite{kennedy2006swarm},
among others. These population-based algorithms share the common characteristic
of using an initial set of random candidate solutions that are later used to
generate a new set of candidates, using a nature-inspired heuristic. Popular EAs
are Genetic Algorithms (GAs), Genetic Programming (GP), grey wolf optimization
(GWO) and Differential Evolution (DE), while examples of (SI) are particle swarm
optimization (PSO) and ant colony algorithms (ACO).

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ  
%OK - M
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but maintain synchronism with
other operations,  all operations need to finish to continue to the next step.
For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. In the previous, the parent operation can continue to the next
iteration, even if a worker has not finished its work.  Several works have
proposed asynchronous algorithms, reporting benefits in execution time, and
scalability. In these works, sub-populations
are processed independently and communicate asynchronously through a central
repository or message queue [refs needed to our previous work].

@@

The parallel execution of multi-population algorithms has been the focus of many
works,  going from earlier hardware-based implementations using transputers,
multi-threaded, and multi-core systems. % many citations here - JJ
Recently the trend has been on
exploiting a higher number of processing units by using GPUs, or distributed
systems; techniques used to achieve this include client-server, web-based, map-reduce,  grid, and cloud computing. % many references here too - JJ
Cloud
computing has become the standard way of running enterprise applications, not
only because of the convenience of the pay-as-you-go model or the non-existent
cost of administration, but also because it offers a way of describing the
infrastructure as part of the code, so that it is much easier to reproduce
results and this has been a boon for scientific computing.  However,  cloud computing
has also been evolving, going from simply putting in the cloud old-style monolithic applications, that is,
applications built on a single stack of services that communicate by layers, to
microservice and serverless architectures that favor the parallel and
asynchronous communication of heterogeneous resources; in the process, a new process for designing {\em cloud native applications} has been created, with brand new methodologies and techniques. In these architectures,
services or even functions are seen as single processing points,  departing
from the monolithic and even distributed paradigms, become a loosely coupled collection of 
{\em stateless functions}, that react to events and only come into existence while they 
are doing some processing. % many references more here - JJ

@@

% There is a big gap here where you go from generic considerations
% about cloud computing to the presentation of this method. In the
% middle, there should be motivation for it, what are you trying to
% address precisely, and how you intend to address it. "Approaching
% multi-population algorithms in cloud computing" is not nearly
% enough. There are several points here, and every one of them needs
% to be addressed and justified
% * Using containers as opposed to a serverless architecture or a
% virtual-machine based one.
% * Combining several metaheuristics instead of using a single one.
% * Reducing the amount of parameters.
% There needs to be a motivation for every one, relate them to the
% objectives of the paper (which need to be clarified) and eventually
% prove the contribution of every one of them to the final result - JJ
Taking these factors into account, in this work, we present Swarm, % maybe a less generic name? - JJ
a
multi-container Docker application that reactively processes isolated and
heterogeneous subpopulations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. % Docker compose is not really for production, more for development environments, you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
Docker
Container images host the stateless functions responsible for running algorithms
on subpopulations. Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of sub-population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.

>>>>

In the last decades, nature-inspired optimization algorithms have been succesfully
applied to solve complex real-world problems \cite{yang2014nature}. Algorithms
inspired by natural processes include evolutionary algorithms (EAs)
\cite{back1996evolutionary} and swarm intelligence (SI) \cite{kennedy2006swarm},
among others. These population-based algorithms share the common characteristic
of using an initial set of random candidate solutions that are later used to
generate a new set of candidates, using a nature-inspired heuristic. Popular EAs
are Genetic Algorithms (GAs) \cite{holland1992adaptation,eiben2003genetic}, 
Genetic Programming (GP) \cite{back1996evolutionary}, Grey Wolf Optimization
(GWO) and Differential Evolution (DE), while examples of (SI) are particle swarm
optimization (PSO) and ant colony algorithms (ACO).

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ  
%OK - M
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but maintain synchronism with
other operations,  all operations need to finish to continue to the next step.
For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. In the previous, the parent operation can continue to the next
iteration, even if a worker has not finished its work.  Several works have
proposed asynchronous algorithms \cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time, and
scalability. In other works, sub-populations
are processed independently and communicate asynchronously through a central
repository \cite{sofea:cec2012,agajaj} or message queues \cite{salza2019speed, guervos2018introducing} .

@@

The parallel execution of multi-population algorithms has been the focus of many
works,  going from earlier hardware-based implementations using transputers ,
multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
Recently the trend has been on exploiting a higher number of processing units 
by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 
client-server , web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
voluntary computing \cite{MilkyWay}, and cloud computing 
\cite{GValdez2015,salza2019speed,valenzuela2015implementing,Sherry2012}. % many references here too - JJ
Cloud computing has become the standard way of running enterprise applications, not
only because of the convenience of the pay-as-you-go model or the non-existent
cost of administration, but also because it offers a way of describing the
infrastructure as part of the code, so that it is much easier to reproduce
results and this has been a boon for scientific computing.  However,  cloud computing
has also been evolving, going from simply putting in the cloud old-style monolithic applications, that is,
applications built on a single stack of services that communicate by layers, to
microservice and serverless architectures \cite{varghese2018next} that favor the parallel and
asynchronous communication of heterogeneous resources; in the process, 
a new process for designing {\em cloud native applications} has been created, 
with brand new methodologies and techniques. In these architectures,
services or even functions are seen as single processing points,  departing
from the monolithic and even distributed paradigms, become a loosely coupled collection of 
{\em stateless functions}, that react to events and only come into existence while they 
are doing some processing. % many references more here - JJ

@@

% There is a big gap here where you go from generic considerations
% about cloud computing to the presentation of this method. In the
% middle, there should be motivation for it, what are you trying to
% address precisely, and how you intend to address it. "Approaching
% multi-population algorithms in cloud computing" is not nearly
% enough. There are several points here, and every one of them needs
% to be addressed and justified
% * Using containers as opposed to a serverless architecture or a
% virtual-machine based one.
% * Combining several metaheuristics instead of using a single one.
% * Reducing the amount of parameters.
% There needs to be a motivation for every one, relate them to the
% objectives of the paper (which need to be clarified) and eventually
% prove the contribution of every one of them to the final result - JJ
Taking these factors into account, in this work, we present Swarm, % maybe a less generic name? - JJ
% agree - Mario
a multi-container Docker application that reactively processes isolated and
heterogeneous subpopulations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. 
% Docker compose is not really for production, more for development environments, 
% you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
Docker Container images host the stateless functions responsible for running algorithms
on subpopulations. Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of sub-population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.
